---
title: "Reproducible research assignment"
author: "David García Sabaté"
date: "24/4/2019"
output: 
  html_document: 
    fig_caption: yes
    self_contained: no
    keep_md: true
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)
```

## Programming assignment 1: reproducible research  

### 1. Code for reading in the dataset and/or processing the data

Download the "activity.csv" dataset

```{r import}
zipUrl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
zipFile <- "repdata_data_activity.zip"
download.file(zipUrl, zipFile, mode = "wb")
unzip(zipFile)
```

Now we have tib_observations tibble object, and see a sample of 10 observations

```{r tibble}
dataframe <- read.csv("activity.csv")
tib_observations <- as_tibble(dataframe)
tib_observations$date <- ymd(tib_observations$date)
head(tib_observations, 10)
```

### 2. Histogram of the total number of steps taken each day

```{r histogram}
sum_steps <- tib_observations %>%
  group_by(date) %>%
  summarize(
    steps = sum(steps)
  )

ggplot(sum_steps, aes(date, steps)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 90))
```

### 3. Mean and median number of steps taken each day

We remove de NA values since we assume that there are no observations in that period of time.

Mean:
```{r}
mean(sum_steps$steps, na.rm = T) 
```

Median:
```{r}
median(sum_steps$steps, na.rm = T)
```

### 4. Time series plot of the average number of steps taken

```{r}
ggplot(sum_steps, aes(date, steps)) +
  geom_line() +
  scale_x_date(date_breaks = "1 week") +
  theme(axis.text.x = element_text(angle = 90))
```

### 5. The 5-minute interval that, on average, contains the maximum number of steps
```{r}
five_minute <- tib_observations %>%
  filter(!is.na(steps)) %>%
  group_by(interval) %>%
  summarize(
    average = mean(steps)
  ) %>%
  arrange(desc(average))

  head(five_minute, 1)
```

The interval that, on average, contains de maximum number of steps is 8:35

### 6. Code to describe and show a strategy for imputing missing data

Total number of missing values:  

```{r}
tib_obsNA <- tib_observations %>%
filter(is.na(steps))
  
(count(tib_obsNA) * 100) / count(tib_observations)
```
Total NAs are 2304 observations. They represent 13% of the whole sample.

Strategy for filling in all of the missing values in the dataset is using the mean for that 5-minute interval.

```{r}
#calculate means for every period in a separate table
tib_media_periodos <- tib_observations %>%
  group_by(interval) %>%
  summarize(
    mean = mean(steps, na.rm = T)
  )

#join means table with the main table
tib_medias <- left_join(tib_observations, tib_media_periodos)
tib_medias$mean <- as.integer(tib_medias$mean)
#Replace NA Values
tib_medias$steps[is.na(tib_medias$steps)] <- tib_medias$mean[is.na(tib_medias$steps)]  

# Resulting table
tib_filled <- tib_medias %>%
  select(-mean)

head(tib_filled,10)
```


### 7. Histogram of the total number of steps taken each day after missing values are imputed

```{r histogram2}
tib_filled %>%
  group_by(date) %>%
  summarize(
    steps = sum(steps)
  )

ggplot(tib_filled, aes(date, steps)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 90))
```

### 8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r}
# weekdays -> 2-6, weekends -> 1 or 7
tib_filled_wdays <- tib_filled %>%
  mutate(
    weekday = wday(date, label = T),
    weekday_int = wday(date),
    is_weekday = weekday_int %in% c(2:6) 
  )

tib_filled_wdays_summary <- tib_filled_wdays %>%
  group_by(interval, is_weekday) %>%
  summarize(
    mean = mean(steps)
  ) 

tib_filled_wdays_summary$is_weekday <- str_replace_all(tib_filled_wdays_summary$is_weekday, c("FALSE" = "Weekends", "TRUE" = "Weekdays"))


ggplot(data = tib_filled_wdays_summary) + 
  geom_line(mapping = aes(x = interval, y = mean)) +
  facet_wrap(~ is_weekday, nrow = 2, ncol = 1) + 
  labs(
    title = "Comparision: number of steps (Weekdays V/S weekends)",
    subtitle = "People tend to walk more on weekends"
  )
```

